Prosper Loan Investment Strategy by Mike Chan
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.

#install.packages("ggplot2", dependencies = T) 
#install.packages("knitr", dependencies = T)
#install.packages("dplyr", dependencies = T)

library(ggplot2)
library(knitr)
library(dplyr)

#getwd()
#setwd("C:/Users/Michael/Google Drive/Udacity P4/P4 PROJECT")

```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data

ld <- read.csv('prosperLoanData_filtered.CSV')
ld <- subset(ld, CreditScoreRangeLower = as.numeric(CreditScoreRangeLower))
ld <- subset(ld, !is.na(EstimatedReturn))
ld <- subset(ld, !is.na(IncomeRange))
ld <- subset(ld, !is.na(DebtToIncomeRatio))
ld$TotalAmountReturn <- with(ld,EstimatedReturn*LoanOriginalAmount)
ld$IncomeRange <- factor(ld$IncomeRange, levels = c("$0","$1-24,999",
                                                    "$25,000-49,999",
                                                    "$50,000-74,999",
                                                    "$75,000-99,999",
                                                    "$100,000+"))
ld$ProsperRating <- factor(ld$ProsperRating, 
                           levels = c("AA","A","B", "C", "D", "E","HR"))

```

# Univariate Plots Section

I would like to display the data of all my data set and what it is like...
```{r echo=FALSE, Univariate_Plots}
summary(ld)
```

Comment: General idea of what to expect of the data set

I would like to see how much investors have earned for using the Prosper 
Loan system...
```{r echo=FALSE, Univariate_Plots1}
qplot(data = ld, TotalAmountReturn, 
      main = "Count of Total Returns for Investors",
      xlim = c(-1000,4000), bins=200)

```

Comment: Investors have nearly received on average, $828 in return per loan  
Investors can possibly earn as high as $4,000.  Interestingly, I'm seeing a 
large spike of investors earning $500.  Notice though, that investors still 
can lose money as we see negative returns.


I would like to show the count of all loans of different prosper ratings...
```{r echo=FALSE, Univariate_Plots2}
qplot(data = ld, ProsperRating, 
      main = "Count of All Loans with Different Prosper Ratings")
```

Comment: Most loans have a prosper rating of C


I would like a count of all loans with different estimated returns...
```{r echo=FALSE, Univariate_Plots3}
qplot(data = ld, EstimatedReturn, main = "Count of All Loans with Different 
      Estimated Returns", xlim = c(0,0.2), bins = 100)
```

Comment: The average estimated return looks to be around 0.09



I would like a count of al loans with different numbers of borrower APRs...
```{r echo=FALSE, Univariate_Plots4}
qplot(data = ld, BorrowerAPR, main = "Count of All Loans with Different 
      Borrower's APR", bins=100, xlim = c(0.05,0.425))
```

Comment: The average normal distribution looks to be at mean of 0.2.  
There is however a sudden big spike at around 0.36 APR.


I would like a count of all loans with different borrower's credit score..
```{r echo=FALSE, Univariate_Plots5}
qplot(data = ld, CreditScoreRangeLower, main ="Count of All Loans 
      with Different Borrower's Credit Score (Lowest Range)")
```

Comment: The avearge Credit Score Range is about 675. 


I would like a count of loan of all different amounts borrowed...
```{r echo=FALSE, Univariate_Plots6}
qplot(data = ld, LoanOriginalAmount, main ="Count of All Loans with 
      Different Amount of Loan Borrowed", bins = 100)
```

Comment: The graph looks skewed to the left.  So the average amount borrowers 
borrow is $8,000.  I notice that borrows tend to borrow in sets of 500s and 
1000s because we can see big spikes at 4,000, 5000, 10000, 15000, 20000, etc.



I would like a count of all loans with different debt to income ratio...
```{r echo=FALSE, Univariate_Plots7}
qplot(data = ld, DebtToIncomeRatio, main = "Count of All Loans with Different 
      Debt to Income Ratio", bins = 100)
qplot(data=ld, DebtToIncomeRatio, main = "Count of All Loans with Different 
      Debt to Income Ratio", bins = 100, xlim = c(0,1.25))
```

Comment: First, I want to point out the big outlier at the Debt to Income Ratio 
of 10, and anything in between around 1.2 to 10 looks to absolutely nothing.  
I then removed the bias by changing my x limit to between 0 and 1.25.  
I see that the average Debt to Income Ratio is about 0.2 .  I also notice the 
large spikes in which my best guess is that borrowers just put in random 
numbers like 0.1, 0.2, etc.

In the next graph we will use log transform to see it more normally distributed



I would like a count of all loans with different debt to income ratio of log 
10 because it looked very skewed before...
```{r echo=FALSE, Univariate_Plots8}
qplot(data = ld, log10(DebtToIncomeRatio+1), main ="Count of All Loans with 
      Different Debt to Income Ratio of log 10", bins = 100, xlim = c(0,0.3))
```

Comment: The average debt to income ratio appears to be around log of 0.09 but 
as the ratio becomes larger, the value of that ratio becomes more continuous 
( Left half of the graph has more break lines than right half).



I would like a count of all loans with different loan statuses...
```{r echo=FALSE, Univariate_Plots9}
qplot(data = ld, LoanStatus, main = "Count of All Loans with Different Loan 
      Statuses")+ theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Comment: Most loans are currently going on and a majority of them are completed.


I would like a count of all loans with different borrower's income range...
```{r echo=FALSE, Univariate_Plots10}
qplot(data = ld, IncomeRange, main = "Count of All Loans with Different 
      Borrower's Income Range") + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Comment: Most borrowers are between the income range of $50,000 and $74,999
I was hoping to see borrowers of income between $1 and $24,999 to be the 
highest because, they need to borrow more.


I would like a count of all loans with different terms
```{r echo=FALSE, Univariate_Plots11}
ld_term <- group_by(ld, Term)
count(ld_term, Term)
```

Comment: A lot of people borrow a loan for 36 months.



# Univariate Analysis

### What is the structure of your dataset?

There are 11,3937 prosper loans with 10 different features.

The PowerPoint Slide "Explanation of the Data Set" explains why I chose the 
features and explaining in depth.

I stated what I found through different codes.

### What is/are the main feature(s) of interest in your dataset?

The main features in the data set are

EstimatedReturn: Because as Investors, we want to use the data to find the 
best return for our investment.

ProsperRating: And to determine what may be the way to invest for a return 
on our loans, the risk of a loan may have the most influence on a 
return out of all features.

### What other features in the dataset do you think will help support your 
### investigation into your feature(s) of interest?

All the 10 features I picked of the 81 different features on slide 
"Explanation of the Data Set" will support my investigation.  
I explain in detail why I chose those features.

### Did you create any new variables from existing variables in the dataset?

I created the total amount returned variable  ("TotalAmountReturn") 
using the code:

ld$TotalAmountReturn <- with(ld,EstimatedReturn*LoanOriginalAmount)

This function helps us to determine what an investor can expect to be returned 
from investing in that loan.  We're not only interested in percentage 
but the amount investors will receive.

### Of the features you investigated, were there any unusual distributions? 
###Did you perform any operations on the data to tidy, adjust, or change the 
###form of the data? If so, why did you do this?

Yes, the count for each Debt To Income ratio was greatly skewed to the left 
and I wanted to make it more normally distributed.  

# Bivariate Plots Section


I would like to see a box plot of all estimated return for each loans of 
different prosper ratings...
```{r echo=FALSE, Bivariate_Plots}
ggplot(data = ld, aes(y = EstimatedReturn, x = ProsperRating)) + 
  geom_boxplot() + ggtitle("Estimated Return per Prosper Rating")
```

Comment:The average estimated return appears to increase as the Prosper Rating 
Risk increases.  However, note the tails from one end changes amongst different 
risks.  For example, we see more tails in "A" and they are above the 
3rd Quantile and we see more tails in "HR" much below the 1st Quantile.


I would like a count of all estimated returns of differnet prosper ratings...
```{r echo=FALSE, Bivariate_Plots1}
qplot(data = ld, EstimatedReturn, fill = ProsperRating, xlim = c(0,0.2), 
      bins = 100, main = "Estimated Return for Certain Prosper Loans 
      of Different Risks")
```

Comment:That we see a pattern.  As the higher higher the estimated return, 
the higher the prosper rating risk showing up.


I would like to see the estimated returns of all borrowers of different 
credit score ranges...
```{r echo=FALSE, Bivariate_Plots2}
ld$CreditScoreRangeLower <- as.character(ld$CreditScoreRangeLower)

ggplot(data = ld, aes(x=CreditScoreRangeLower, y=EstimatedReturn)) + 
    geom_boxplot() +
    ggtitle("Estimated Return for All Credit Score Ranges")

ld$CreditScoreRangeLower <- as.integer(ld$CreditScoreRangeLower)
```

Comment: The investor's estimated return decreases as borrower's credit score 
increases.
The Credit Score Range also follows the Prosper Rating.  
The lower the Credit Score, the higher the prosper rating score risk 
tends to be.

I would like to see how the typical borrower's APR for each loan
```{r echo=FALSE, Multivariate_Plots6}

ggplot(ld, aes(x= ProsperRating, y=BorrowerAPR)) + 
  geom_boxplot() +
  ggtitle("Borrower's APR per Prosper Rating")

```

Comment:This box plot is a decision making tool for investors to invest in a certain APR in each Prosper Rating for greater gains.
For example, in "C" rated loans, investor should only invest in those 
loans when Borrower APR is between 0.21 and 0.24.  

I would like to see a boxplot of different estimated losses per each loan 
of different prosper ratings...
```{r echo=FALSE, Bivariate_Plots3}
ggplot(data = ld, aes(y = EstimatedLoss, x = ProsperRating)) + geom_boxplot() +
    ylim(0,0.2) + ggtitle("Estimated Loss per Prosper Rating")
```

Comment: The higher one's borrower rating is, the higher the estimated loss 
(unrecoverable money through charge offs) will be for us as investors


I would like to see a boxplot of different estimated losses for each loan status..
```{r echo=FALSE, Bivariate_Plots4}
ggplot(data = ld, aes(y = EstimatedLoss, x = LoanStatus)) + geom_boxplot() +
  ylim(0,0.2) + ggtitle("Estimated Loss per Loan Status") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  geom_hline(yintercept = 0.11 , color = "red") + 
  geom_hline(yintercept = 0.075 , color = "green")
```

Comment:Loans that are charged off, defaulted, or past due usually have 
higher estimated losses to the investor
Loans that are completed, current, and in the final payment progress are loans 
with less estimated losses


I would like to see a scatterplot of all different estimated returns of all 
different loan amounts... 
```{r echo=FALSE, Bivariate_Plots5}
ggplot(data = ld, aes(y = EstimatedReturn , x = LoanOriginalAmount)) + 
  geom_jitter(alpha = 0.1) + stat_smooth(color = "red") + 
  ggtitle("Estimated Return for Typical Loan Amounts")
```

Comment: There is too much noise to conclude that the balance of the loan 
has any impact on the estimated return.


I would like to see a scatter plot of all different esimated returns from 
all different borrower APR.
```{r echo=FALSE, Bivariate_Plots6}
ggplot(data = ld, aes(y = EstimatedReturn , x = BorrowerAPR)) + 
  geom_jitter(alpha = 0.05) + ylim(0,0.2) + stat_smooth(color = "red") + 
  ggtitle("Estimated Returns for Every Borrower APR")
```

Comment:The estimated return does increase as the BorrowerAPR increases but 
variation also increases as well.  


I would like to see a box plot of different credit score ranges of different 
loan statuses...
```{r echo=FALSE, Bivariate_Plots7}
ggplot(ld,aes(x = LoanStatus, y = CreditScoreRangeLower)) + 
  geom_boxplot(lwd = 1) + ggtitle("Credit Score Range per Loan Status") + 
  geom_hline(yintercept = 700, color = "red") + 
  geom_hline(yintercept = 680, color = "blue")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Comment:loans that were cancelled, chargedoff, or defaulted usually are 
borrowers with lower average Credit Scores to begin with.  
Loans that are currently going on, completed, or in the final payment 
progress are loans with borrowers whose credit scores are much higher.  
The summary here is that the lower one's credit score, the higher the chance 
of having a borrower not being able to finish paying off their loans 
(or not receiving in the first place).

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the 
###investigation. How did the feature(s) of interest vary with other features 
###in the dataset?

The higher the Prosper Rating, the more likely higher the return 
(but more volatile)

The higher one's Credit Score, the higher the Prosper Rating.

The lower the Credit Score Range, the lower the estimated return

The higher the Prosper Rating, the higher the estimated loss percentage.

The higher a borrower's APR, the higher the investor's individual return 
(but more volatile).

Loans that have been completed, currently going , and payments in final 
progress are loans in which borrowers have higher credit score than loans 
that have been canceled, charged off, and defaulted.

### Did you observe any interesting relationships between the other features 
###(not the main feature(s) of interest)?

I notice that as the Prosper Rating risk increases, the spread for estimated 
return and loss increases.  This makes sense because the higher the risk, the 
more likely a borrower can fail to repay back the loan thus resulting in charge 
off or default, and thus resulting in investors losing money.

### What was the strongest relationship you found?

The strong relationship I found usually involves between risks and rewards, 
the higher the risk of a borrower (lower credit core and higher prosper rating)
, the higher the return but also higher the loss.

# Multivariate Plots Section


I would like to see the percentage of counts of loan statuses per every 
prosper rating loans...
```{r echo=FALSE, Multivariate_Plots}
ggplot(ld, aes(ProsperRating)) + 
  geom_bar(aes(fill = LoanStatus), position = 'fill') + 
  ggtitle("Percentage of Loan Statuses Per Prosper Rating")
```

Comment:The higher the loan risk, the more likely the loan will end up being 
charged off and defaulted.  And also more likely borrowers will fail to pay on 
time.


I would like to see the percentage of borrower's income range for all 
prosper rating loans...
```{r echo=FALSE, Multivariate_Plots1}
ld<- subset(ld,!is.na(ProsperRating))
ggplot(ld, aes(IncomeRange)) + 
  geom_bar(aes(fill = ProsperRating), position = 'fill') + 
  ggtitle("Percentage of Income Range Per Loan Status") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Comment:The higher the borrower's income, the more likely their loans would 
be lower risk.  The lower one's income range, the more percentage their loans 
would be higher risk.  It's probably because a borrower with in the lower 
income range would most likely have sufficient income to pay back the loan 
thus putting that borrower at a higher risk.


I would like to see the percentage of income range for different loan statuses..
```{r echo=FALSE, Multivariate_Plots2}
ggplot(ld, aes(IncomeRange)) + 
  geom_bar(aes(fill = LoanStatus), position = 'fill') + 
  ggtitle("Percentage of Income Range Per Prosper Rating") + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Comment:The lower one's income range, the more likely their loans would be 
charged off, defaulted and have late payments



I would like to see different graphs of different correlations between, 
borrower's credit score, income range, and prosper rating..
```{r echo=FALSE, Multivariate_Plots3}
ggplot(data = ld,aes(CreditScoreRangeLower)) + 
  geom_bar(aes(fill = IncomeRange), position = 'fill') + 
  facet_wrap(~ProsperRating) + 
  ggtitle("Correlation between Credit Range, Income Range, and Prosper Rating") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Comment:The higher the risk of a prosper loan, the more likely the borrower's 
income range will be lower (as we start to see more brown, pink, and brown bars)
and the borrower's credit score will be lower (the bar chart seem to be shifting
to the left as the prosper rating increases).


I would like to see the estimated return of from each loan of different 
borrower's APR and prosper rating...
```{r echo=FALSE, Multivariate_Plots4}
ggplot(aes(x = BorrowerAPR, y = EstimatedReturn), data = ld) + 
  geom_point(aes(color=ProsperRating)) + ylim(-0.2,0.3) + xlim(0.05,0.425) + 
  ggtitle("Estimated Return per Borrower APR per each Prosper Loan Risk")
```

Comment:We see that different prosper ratings are directly stacked on top of 
each other.  We can use this data as an advantage for investors to decide on 
which loans for optimal returns. 


I would like to see an average moving line through different estimated returns 
through borrower's APR and prosper rating loans...
```{r echo=FALSE, Multivariate_Plots5}
ggplot(aes(x = BorrowerAPR, y = EstimatedReturn), data = ld)+
  geom_point(alpha = 0.5, color="grey") + 
  stat_smooth(aes(color=ProsperRating)) +
  ylim(-0.2,0.3) + xlim(0.05,0.425) + 
  ggtitle("Estimated Return per BorrowerAPR per Prosper Rating Loan")
```

Comment:That we see the averages of each ratings do not intersect one another. 
Therefore, I see a pattern..


I would like to see an average line through all loans of different
estimated returns of different borrowers APR with different prosper rating...
```{r echo=FALSE, Multivariate_Plots7}
ggplot(aes(x = CreditScoreRangeLower, y = EstimatedReturn), data=ld) +
  geom_point(color="grey") + ylim(-0.1,0.25) + xlim(600,900) + 
  stat_smooth(aes(color = ProsperRating)) + 
  ggtitle("Estimated Return for Each Credit Score")
```

Comment:It's difficult to determine which Prosper Rating to invest at different 
Credit Scores.  The variance is also too high and other factors are more 
important to contribute the decision.

I would like to see estimated returns for each credit score per prosper
rating...
```{r echo=FALSE, Multivariate_Plots9}
ggplot(data = ld,aes(x = CreditScoreRangeLower,y = BorrowerAPR)) +
  geom_point(alpha = 0.005) + 
  ggtitle("Estimated Return for Each Credit Score per Prosper Rating") + 
  stat_smooth(aes(color=ProsperRating))
```

Comment:There appears to be a consistent horizontal pattern for each Prosper 
Rating.  By consistency, the higher the rating of a loan, the higher and 
consistent the borrower's APR.



I would like to see total amount returned (estimated return times loan 
original amount) for each prosper rating...
```{r echo=FALSE, Multivariate_Plots10}
qplot(data = ld,EstimatedReturn*LoanOriginalAmount, fill = ProsperRating, 
      xlim = c(-100,4000), binwidth=50) + 
  ggtitle("Estimated Return X Loan Original Amount per Prosper Rating")
```

Comment:At most, an investor can make almost $4000 on a loan.  We see that an 
invstor investing in HR can struggle to make more than $1000 in return.



I would like to see investor's total return (estimated return times 
loan original amount) for each term...
```{r echo=FALSE, Multivariate_Plots11}
ld <- group_by(ld,Term = as.character(Term))

qplot(data = ld,EstimatedReturn*LoanOriginalAmount, fill = Term, 
      xlim = c(-100,3500), binwidth=50) + 
  ggtitle("Estimated Return x Loan Original Amount per Term")
```

Comment: Investing in longer term can yield more earnings to the investor



I would like to see all investor's total return (estimated return times 
loan original amount) at each prosper rating at different credit score...
```{r echo=FALSE, Multivariate_Plots12}

qplot(data = ld,x = EstimatedReturn*LoanOriginalAmount, 
      fill = CreditScoreRangeLower, xlim = c(-1000,3000)) + 
  facet_wrap(~ProsperRating) + 
  ggtitle("Estimated Return x Loan Original Amount per Credit Score")


ld$CreditScoreRangeLower <- as.integer(ld$CreditScoreRangeLower)
```

Comment:An investor on average can achieve the highest return from loan "E".  
Loan "E" also has appears to have the least variance so it makes achieving 
that higher return much easier for the investor.
Investors can benefit earnings from borrowers with higher Credit Scores in
safer loans: "AA" and "A" where we see some slight "pink and purple".  



I would like to see all investor's total return (estimated return times 
loan original amount) at each prosper rating at different loan terms...
```{r echo=FALSE, Multivariate_Plots13}
qplot(data = ld,x = EstimatedReturn*LoanOriginalAmount, 
      fill = Term, binwidth = 50, xlim = c(-100,3000)) + 
  facet_wrap(~ProsperRating) + 
  ggtitle("Estimated Return x Loan Original Amount per Proper Rating")
```

Comment:An Investor will not see the benefit of a loan borrowed for 60 months
at "HR" (as we cannot see any blue spots in that graph). 
However, an investor may benefit greatly from that "HR" loan because of the 
spike we see returning almost $25,00.
We can hardly see any great gains from 12 month term loans.



I would like to see estimated returns for each prosper rating..
```{r echo=FALSE, Multivariate_Plots14}
qplot(data = ld, EstimatedReturn, fill = ProsperRating, 
      geom = "density", xlim = c(0,0.2), alpha=0.1)
```

Comment: Indeed, the average estimated return does increase as the risk 
of the Prosper rating increases.  
However, a loan of "D" or "E" can achieve higher returns (at and above 0.15)
than loan "HR".  Estimated Returns for "HR" do not appear pass 0.13.



I would like to see a density chart of all investor's total estimated return
for each prosper rating...
```{r echo=FALSE, Multivariate_Plots15}
qplot(data = ld, EstimatedReturn*LoanOriginalAmount, fill = ProsperRating, 
      alpha = 0.05, geom = "density", 
      main = "Occurrences of Total Estimated Returns") + 
  scale_x_continuous(limits = c(-100,3000), breaks = seq(0,3000,500)) + 
  geom_vline(xintercept = 872.63)
```

Comment:The average investor return per loan is $872.63.  It looks like
that can most likely be achieved with a "AA" or "B" loan. 
Investing in the "HR" loan, it is very unlikely an investor can get an
investment of $872.63.  In fact it looks to be the average return of
the "HR" loan is $500


I would like to see a density chart of all investor's total amount 
borrowered with APR for each loans of different prosper rating...
```{r echo=FALSE, Multivariate_Plots16}
qplot(data = ld,(1 + BorrowerAPR) * LoanOriginalAmount, alpha = 0.05,
      fill = ProsperRating,  geom = "density", 
      main ="Occurrences of Total Amount Borrowed with APR")
```

Comment: The lower the risk rating of the borrower, the more the 
borrower will borrow.  The maximum amount for borrowers to borrow on 
loans rated "HR" appears to be at about 1,000, then the risk levels of 
Prosper decreases as borrowers seek to borrow more.

Probably the reason why it doesn't look like borrowers of loan HR can 
borrow more than 1,000 is because of high risk of Prosper Loan company to
not want to lose.  You obviously do not want to lend out a whole lot of money
to risky people and have a high chance of not be able to have them pay back 
fully.  This so on for the other less riskier loans.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the 
###investigation. Were there features that strengthened each other in terms of 
###looking at your feature(s) of interest?

There are certain percentage of Borrower's APR in which the investor should 
not invest in because the previous less risky loans will tend to have a better
return than.

The longer the term, the higher the estimated return.

Sometimes, it is not worth investing in an HR Loan between in an APR less than 
0.3 when a much return can be invested in other less riskier loans less than 
the APR of 0.2.

### Were there any interesting or surprising interactions between features?

Yes on average an investor who invests in the "HR" loan will have high return 
but the highest return can be achieved through E or D loan.  
It is low probability an investor can achieve that sort of return in the 
upper quartiles of loans E and D but it is something that mostly unlikely to 
be achieved through loan HR.  

### OPTIONAL: Did you create any models with your dataset? Discuss the 
###strengths and limitations of your model.

N/A

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

ggplot(data = ld,aes(CreditScoreRangeLower)) + 
  geom_bar(aes(fill = IncomeRange), position = 'fill') + 
  facet_wrap(~ProsperRating)  +
  ggtitle("Correlation between Credit Range, Income Range, and Prosper Rating") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

### Description One

Looking at the graphs below starting with the lowest risk ("AA") and 
examining up to the higher risk ("HR").   A few observations.
We start to see more colors of the lower income ranges (see more red, 
brown, and pink emerging)

We start to see the graph shifting to the left 
(meaning the borrower's credit score decreases)

This concludes that a prosper loan is most likely determined by a 
borrower's income range and credit score.

The lower a borrower's income range, the less likely that borrower would 
have sufficient balance to pay back, thus higher the Prosper Rating would be.

The lower a borrower's credit score, the less likely the borrower would have 
the ability to pay back the loan.  Most likely because the borrower doesn't 
have a stable job or have bad habits of making late payments.

This is important to the investor to understand why a prosper loan risk rating 
is rated the way it is and the variables that may contribute to that rating 
causing investors to either have gains or losses whe investing.

### Plot Two
```{r echo=FALSE, Plot_Two}
ggplot(data = ld, aes(y = EstimatedReturn, x = ProsperRating)) + 
  geom_boxplot() + ggtitle("Estimated Return per Prosper Rating")
```

### Description Two

We always tend to think that the higher the risk of a loan, the higher chance 
on average we will receive a higher estimated return than the other lower risk
loans.  However, that is not the case by looking at loan HR and comparing that
with loans E and D.  The average estimated return for E and D look to be 
identical to the average return for loan HR.  And also, looking at the upper
quartiles of loans E, D, and HR, one can achieve a highest return on loans E 
and D that may unlikely to be achieved on loan HR.  Also HR has a lot of 
potential risk of losses for investors.  For this,  an investor may be better
off looking to invest in loans E and D for higher returns and ignoring 
investing at all in HR if possible.

### Plot Three
```{r echo=FALSE, Plot_Three}

ggplot(ld, aes(x= ProsperRating, y=BorrowerAPR)) + 
  geom_boxplot() + ggtitle("Borrower's APR per Prosper Rating")

```

### Description Three

This plot helps investors decide on which typical Prosper Rating loan to invest in based on the borrower's APR.  Investor should target the filled color boxes and 
avoid the "tail ends" if possible (the plots below the lower quartile and the 
plots above the upper quartile). 

For example, in loan B, an investor should invest in that loan when the APR is
between 0.17 and 0.19.  Anything less than an APR 0.17, an investor would be 
better off investing in loan A and anything higher than 0.19 an investor 
should be better off investing in loan C.

------

# Reflection

### Where did I run into difficulties in the Analysis?

One of the great difficulties I had was before my analysis was greatly trying
to understand the purpose of each feature.  With the 81 different features, 
it's difficult to pick which ones would offer great relationships between each
other and insights to what investors and borrowers want.  I was surprised that 
an investor debt to income ratio did not result in much correlation to any 
features.  I was hoping the more debt/income ratio a borrower have, the more
likely he/she will struggle to pay it back thus resulting in higher prosper 
loan rating.

### Where did I find success?

I found success after plotting bivariate and multivariate charts thus finding 
patterns between certain features.  Some of the features I chose were ones that
would give investors good information as to how they would choose where to 
invest.  And after doing a little bit of exploring between data of each 
desired feature, that's where I found success.
I succeeded knowing for investors that loan HR should not be invested at all. 
I found this out by plotting graphs of the trends between Borrower APR, 
Prosper Rating, and the Estimated Return.  As an investor, we would usually 
want information such as the borrower's credit score, income range, and APR of
a loan.  Those 3 key information would allow investors see if investing in that
certain loan would probably have good returns/losses.
In general, I found success by thinking what an investor would think: What
are we most concerned about from a given loan?  And we want to know about the 
Borrower's history as much as possible to minimize risk and maximize returns.

### How could the analysis be enriched in future work?

Most of the data collected are loans currently being paid off slowly by the 
investor.  It's better to revisit the data after all current loans have ended
in either completed or incomplete (defaulted or charged off).  In the end, 
what matters to an investor is a completely paid off loan will always result 
in a return for the investor and an incomplete loan will always result in some 
loss to the investor.  From the looks of the count of complete or incomplete 
loans, it looks like most loans in the end will be completed.

What would be more enriching is to get to know about the investor's situation.
Of course an investor would want to the best possible return but there are 
factors like how soon an investor would like all his money receive back all
his money including earnings. Therefore, we should analyze with the term being
12 months if an investor wants everything returned as soon as possible.  Or 
what if the investor is very emotional and anxious about losses, then we should
probably calculate the best possible return from maybe loans A, B, and C 
even though loans D and E offer the highest average return.  These are 
variables that needs to be taken into account to see how we can best 
diversify the choosing of the loans for best returns suitable for each 
investor's unique situation.


